The \dealii{} Library, Version 9.0

### 2018-04-23 9.0/paper.tex
Update the section on tutorial/code gallery programs.

The major changes of this release are:
\begin{itemize}
\item xxxx something with manifolds xxx
\item Support for particle-in-cell methods
\item Dedicated support for automatic and symbolic differentiation.
\item Interfaces to more external libraries and programs.
\item C++11 is now both required and used.
\item Support for GPU computations.

\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

@@

\item New python tutorial program tutorial-1; as well as
  updates to step-37. In addition, the separate code
  gallery of \dealii{} has gained a number of new entries.

  \item More than 330 other features and bugfixes.
\end{itemize}
The more important ones of these changes will be detailed in the
following section.  Information on how to cite \dealii{} is provided
in Section \ref{sec:cite}.

@@

@@

While \dealii{} is a package intended to solve problems with the
finite element method -- i.e., using continuous or discontinuous
\textit{fields} --, is often convenient in fluid dynamics problems to
couple the continuum description of phenomena with particles. These
particles, advected along with the numerical approximation of the flow
field, are then either used to visualize properties of the flow, or to
advect material properties such as the viscosity of inhomogenous
mixtures of fluids. If each particle is associated with the cells
of a mesh, these methods are often referred to as particle-in-cell (PIC).

@@

A much more detailed view of the underlying algorithms can be found in
\cite{GLHPB18}. A longer report is at \cite{GHPB16}.

@@

A dedicated differentiation module has been created to provide support for
automatic and symbolic differentiation. Automatic differentiation is a set of
technique to evaluate the derivatives of a function defined by a compute
program. Currently the Adol-C and Sacado libraries are supported through a
unified interface. In practice, this support means that Adol-c and Sacado
data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
to represent degree-of-freedom values.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and updated tutorial programs}

\marginpar{OLD}
In addition to the updated tutorial programs mentioned in the previous
section, this release of \dealii{} includes three new tutorials TODO: old:
\begin{itemize}
 \item {\bf step-55} explains how to solve the Stokes
 equations efficiently in parallel. It is a good introduction to solving
 systems of PDEs in parallel, discusses optimal block
 preconditioners, and demonstrates other aspects like error computation.
 Inverses of individual blocks of the linear system are approximated with an algebraic
 multigrid preconditioner.

 \item {\bf step-56} shows how to apply geometric multigrid
 preconditioners on a subset of a system of PDEs. The problem solved here is the
 Stokes equations, like in step-55.

 \item {\bf step-57} solves the stationary Navier-Stokes equations.
 The nonlinear system is solved using Newton's method on a sequence of adaptively refined
 grids. The preconditioner is again built on a block factorization of the saddle point
 system like in step-55 and step-56, but the non-symmetric terms stemming from the
 nonlinear convective part requires more sophisticated solvers. The benchmark problem,
 flow in the 2d lid-driven cavity, requires a continuation method for high Reynolds
 numbers.
 \end{itemize}

In addition to tutorials, \dealii{} has a separate ``code gallery'' that
consists of programs shared by users as examples of what can be
done with \dealii{}. While not part of the release process, it is nonetheless
worth mentioning that the set of new programs since the last release covers
the following topics:
\marginpar{OLD}
  \begin{itemize}
    \item Quasi-static quasi-incompressible visco-elastic material behavior;
    \item Multiphase Navier-Stokes flow;
    \item The evolution of global-scale topography on planetary bodies;
    \item Goal-oriented elastoplasticity.
  \end{itemize}

>>>>

The major changes of this release are:
\begin{itemize}
\item xxxx something with manifolds xxx
\item Support for particle-in-cell methods
\item Dedicated support for automatic and symbolic differentiation.
\item Interfaces to more external libraries and programs.
\item C++11 is now both required and used.
\item Support for GPU computations.
\end{itemize}
These will all be discussed in more detail in the
following section. In addition, this release contains the following changes:
\begin{itemize}
\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

@@

\item A new python tutorial program tutorial-1; as well as
  updates to step-37. In addition, the separate code
  gallery of \dealii{} has gained a number of new entries.
\end{itemize}
Beyond these changes, the changelog lists more than 330 other features and bugfixes.

@@

\marginpar{All: please edit as appropriate}

(i) separation between manifold and boundary ids, (ii) manifold
smoothing via the transfinite mapping, (iii) maybe: grid generator functions already attach manifolds?

Previous text:

@@

\marginpar{Rene: Please edit as appropriate}
While \dealii{} is a package intended to solve problems with the
finite element method -- i.e., using continuous or discontinuous
\textit{fields} --, is often convenient in fluid dynamics problems to
couple the continuum description of phenomena with particles. These
particles, advected along with the numerical approximation of the flow
field, are then either used to visualize properties of the flow, or to
advect material properties such as the viscosity of inhomogenous
mixtures of fluids. If each particle is associated with the cells
of a mesh, these methods are often referred to as particle-in-cell (PIC).

@@

A much more detailed view of the underlying algorithms can be found in
\cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
implementation here originated in the \aspect{} code, see \cite{KHB12,HDGB17}.

@@

\marginpar{J-P: Please edit as appropriate}
Automatic differentiation is often used to automatically derive
residuals from the stored energy functional, and to derive Jacobian
matrices from residual vectors for simulations that use complicated
material models. Examples can be found widely for nonlinear solid
mechanics as well as for nonlinear viscosity models in fluid flow.

\dealii{} has had a tutorial program (step-33) since 2007 that
demonstrates this technique based on the Trilinos Sacado package, but
the functionality was not available pervasively throughout
deal.II. This has changed with release 9.0 where support for
differentiation has been collected in one module that provides both
automatic and symbolic differentiation. Currently, this module relies
on either the Adol-C and Sacado libraries, through a unified
interface. In practice, this support means that Adol-C and Sacado data
types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
and related classes that are generally used to assemble linear systems
and right hand sides.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tutorial and code gallery programs}

This release does not contain any new tutorial programs, though
several have been updated extensively for the changes to the manifold
handling as well as to adjust for current functionality
compared to what that was available when the programs were first written.

\dealii{} has a separate ``code gallery'' that
consists of programs shared by users as examples of what can be
done with \dealii{}. While not part of the release process, it is nonetheless
worth mentioning that the set of new programs since the last release covers
the following topics:
\begin{itemize}
\item The multipoint flux mixed finite element method (MFMFE) applied
  to the Darcy problem of porous media flow;
\item A linearized active skeletal muscle model with application to
  the simulation concentric contraction of the human biceps brachii;
\item A parallel implementation of the Local Discontinuous Galerkin
  (LDG) method applied to the Poisson equation.
\end{itemize}
With these additions, the code gallery now contains 10 different applications.
